---
title: "More than noise?: Make single trial DF signal variability measures"
author: "José C. García Alanis"
date:  "`r Sys.Date()`"
output:
  html_document:
    theme: lumen
    highlight: pygments
    css: "styler.css"
---

Created with `r R.version.string`

```{r, setup, include = FALSE}
knitr::opts_chunk$set(class.source = "customCode",
                      class.output = "customOutput",
                      class.message = "customWarning")
if (interactive()) {
  path <- dirname(rstudioapi::getSourceEditorContext()$path)
} else {
  path <- normalizePath('./')
}
setwd(path)
source('utils.R')
```

```{r session-preferences, include = FALSE}
load.package(
        c('rjson', 'dplyr', 'tidyr', 'kableExtra', 'stringr')
)
```

# 1) Import the data

## 1.1) Odd/Even task

### 1.1.1) Amplitude variability

```{r import-data-amplitude-var-oddeven}
# get paths to Odd/Even task PSD
paths <- fromJSON(file = "paths.json")
fpath_signal_var_oddeven_amp <- paste(
        paths$bids,
        'derivatives',
        'signal_variability',
        'oddeven_*',
        '*_amp_*single_trial.tsv',
        sep = '/'
)

# gather files matching pattern
signal_variability_files_oddeven_amp <- Sys.glob(fpath_signal_var_oddeven_amp)

# import the data
signal_variability_list_oddeven_amp <- lapply(
        signal_variability_files_oddeven_amp,
        read.table,
        sep = '\t',
        header = TRUE,
        colClasses = c(
                rep('integer', 3),
                rep('character', 3),
                'integer',
                'numeric',
                rep('numeric', 11)
        )
)

# row bind RT data.frames
signal_variability_df_oddeven_amp <- bind_rows(
        signal_variability_list_oddeven_amp,
        .id = NULL
)
rm(fpath_signal_var_oddeven_amp,
   signal_variability_files_oddeven_amp,
   signal_variability_list_oddeven_amp
)

# rename columns
names(signal_variability_df_oddeven_amp)[11:16] <- c(
        "multiscale_entropy", "ms_1", "ms_2", "ms_3", "ms_4", "ms_slope" )
```

### 1.1.2) Frequency variability

```{r import-data-frequency-var-oddeven}
# get paths to Odd/Even task PSD
paths <- fromJSON(file = "paths.json")
fpath_signal_var_oddeven_freq <- paste(
        paths$bids,
        'derivatives',
        'signal_variability',
        'oddeven_*',
        '*_freq_*single_trial.tsv',
        sep = '/'
)

# gather files matching pattern
signal_variability_files_oddeven_freq <- Sys.glob(fpath_signal_var_oddeven_freq)

# import the data
signal_variability_list_oddeven_freq <- lapply(
        signal_variability_files_oddeven_freq,
        read.table,
        sep = '\t',
        header = TRUE,
        colClasses = c(
                rep('integer', 3),
                rep('character', 3),
                'integer',
                'numeric',
                rep('numeric', 3)
        )
)

# row bind RT data.frames
signal_variability_df_oddeven_freq <- bind_rows(
        signal_variability_list_oddeven_freq,
        .id = NULL
)
rm(fpath_signal_var_oddeven_freq,
   signal_variability_files_oddeven_freq,
   signal_variability_list_oddeven_freq
)

# rename cols
names(signal_variability_df_oddeven_freq)[9:10] <- c("1f_offset", "1f_exponent")
```

### 1.1.3) Combine Odd/Even data

```{r join-amp-freq-tables-oddeven}
signal_variability_oddeven <- signal_variability_df_oddeven_amp %>%
        left_join(., signal_variability_df_oddeven_freq,
                  by = c("subject", "epoch", "sensor", "condition",
                         "stimulus", "window", "accuracy", "rt")
        ) %>%
        mutate(task = 'Odd/Even')

# tidy up
rm(signal_variability_df_oddeven_amp, signal_variability_df_oddeven_freq)
```

## 1.2) Number/Letter task

## 1.2.1) Amplitude variability

```{r import-data-amplitude-var-numberletter}
# get paths to Odd/Even task PSD
paths <- fromJSON(file = "paths.json")
fpath_signal_var_numberletter_amp <- paste(
        paths$bids,
        'derivatives',
        'signal_variability',
        'numberletter_*',
        '*_amp_*single_trial.tsv',
        sep = '/'
)

# gather files matching pattern
signal_variability_files_numberletter_amp <- Sys.glob(
        fpath_signal_var_numberletter_amp
)

# import the data
signal_variability_list_numberletter_amp <- lapply(
        signal_variability_files_numberletter_amp,
        read.table,
        sep = '\t',
        header = TRUE,
        colClasses = c(
                rep('integer', 3),
                rep('character', 3),
                'integer',
                'numeric',
                rep('numeric', 11)
        )
)

# row bind RT data.frames
signal_variability_df_numberletter_amp <- bind_rows(
        signal_variability_list_numberletter_amp,
        .id = NULL
)
rm(fpath_signal_var_numberletter_amp,
   signal_variability_files_numberletter_amp,
   signal_variability_list_numberletter_amp
)

# rename columns
names(signal_variability_df_numberletter_amp)[11:16] <- c(
        "multiscale_entropy", "ms_1", "ms_2", "ms_3", "ms_4", "ms_slope" )
```

## 1.2.2) Frequency variability

```{r import-data-frequency-var-numberletter}
# get paths to Odd/Even task PSD
paths <- fromJSON(file = "paths.json")
fpath_signal_var_numberletter_freq <- paste(
        paths$bids,
        'derivatives',
        'signal_variability',
        'numberletter_*',
        '*_freq_*single_trial.tsv',
        sep = '/'
)

# gather files matching pattern
signal_variability_files_numberletter_freq <- Sys.glob(
        fpath_signal_var_numberletter_freq
)

# import the data
signal_variability_list_numberletter_freq <- lapply(
        signal_variability_files_numberletter_freq,
        read.table,
        sep = '\t',
        header = TRUE,
        colClasses = c(
                rep('integer', 3),
                rep('character', 3),
                'integer',
                'numeric',
                rep('numeric', 3)
        )
)

# row bind RT data.frames
signal_variability_df_numberletter_freq <- bind_rows(
        signal_variability_list_numberletter_freq,
        .id = NULL
)
rm(fpath_signal_var_numberletter_freq,
   signal_variability_files_numberletter_freq,
   signal_variability_list_numberletter_freq
)

# rename cols
names(signal_variability_df_numberletter_freq)[9:10] <- c(
        "1f_offset", "1f_exponent"
)
```

### 1.2.3) Combine Number/Letter data

```{r join-amp-freq-tables-numberletter}
signal_variability_numberletter <- signal_variability_df_numberletter_amp %>%
        left_join(., signal_variability_df_numberletter_freq,
                  by = c("subject", "epoch", "sensor", "condition",
                         "stimulus", "window", "accuracy", "rt")
        ) %>%
        mutate(task = 'Number/Letter')

# tidy up
rm(signal_variability_df_numberletter_amp, signal_variability_df_numberletter_freq)
```


# 2) Combine Odd/Even and Number/Letter data

```{r rowbind_task_dataframes}
signal_variability_st <- bind_rows(
        signal_variability_oddeven,
        signal_variability_numberletter
)

rm(signal_variability_oddeven, signal_variability_numberletter)
```

# 3) Keep complete cases

Here we look for subject for wich we have data from both tasks/sessions:

```{r complete-subjects}
complete_subjects <- signal_variability_st %>%
        select(subject, task) %>%
        unique() %>%
        arrange(subject) %>%
        mutate(val = 1) %>%
        tidyr::pivot_wider(names_from = task, values_from = val) %>%
        filter(complete.cases(.))
```


Also, we exclude trials with extreme RTs (abs(z scored RT) > 3.0):

```{r exclude-non-complete-and-extreme-RTs}
signal_variability_st <- signal_variability_st %>%
  filter(subject %in% complete_subjects$subject) %>%
  group_by(subject, task, condition) %>%
  mutate(logrt = log(rt)) %>%
  mutate(zrt = (logrt - mean(logrt, na.rm = TRUE)) / sd(logrt, na.rm = TRUE)) %>%
  filter(abs(zrt) < 3.0) %>%
  ungroup() %>%
  mutate(condition = stringr::str_to_title(condition),
         stimulus = stringr::str_to_title(stimulus),
         window = stringr::str_to_title(window)) %>%
  mutate(acc = accuracy) %>%
  mutate(accuracy = ifelse(accuracy == 1, "Correct", "Incorrect")) %>%
  filter(accuracy == 'Correct') %>%
  select(-acc)
```

## 3.1) Check number of trials available for analysis

By subject and task:

```{r trial-numbers}
# compute trial number for analysis
trial_numbers <- signal_variability_st %>%
        select(subject, epoch, task) %>%
        unique() %>%
        group_by(subject, task) %>%
        reframe(n = sum(!is.na(epoch)))

trial_numbers %>%
        kbl() %>%
        kable_paper("hover", full_width = F)
```

On average:

```{r average-trial-numbers}
trial_numbers %>%
        group_by(task) %>%
        summarise(average_n = mean(n)) %>%
        kbl() %>%
        kable_paper("hover", full_width = F)
```

## 3.2) Sample descriptives

```{r sample-descriptives}
fpath_subject_data <- paste(paths$bids,
                            'subject_data.csv',
                            sep = '/')
subject_data <- read.csv(fpath_subject_data, sep = ';')

subject_data <- subject_data %>%
  filter(Subject %in% complete_subjects$subject)

subject_data %>%
  summarise(m_age = mean(age),
            sd_age = sd(age),
            min_age = min(age),
            max_age = max(age)) %>%
  kbl() %>%
  kable_paper("hover", full_width = F)

```

```{r}
subject_data %>%
  group_by(sex) %>%
  count(sex) %>%
  kbl() %>%
  kable_paper("hover", full_width = F)
```


# 4) Save DF

```{r save-single-trial-df}
fpath_single_trial_var <- paste(paths$bids,
                                'derivatives',
                                'analysis_dataframes',
                                'single_trial_variability.rds',
                                sep = '/')
dir.create(dirname(file.path(fpath_single_trial_var)), showWarnings = FALSE)
saveRDS(signal_variability_st,
        file = fpath_single_trial_var)
```
```{r save-measures-in-separate-files}
measures <- c("permutation_entropy", "weighted_permutation_entropy",
  "multiscale_entropy", "ms_1", "ms_2", "ms_3", "ms_4",
  "activity", "mobility", "complexity",
  "1f_offset", "1f_exponent",
  "spectral_entropy")
for (meas in measures) {
  meas_df <- signal_variability_st %>%
    select(subject, epoch, sensor,
           condition, stimulus, window, task, all_of(meas)) %>%
    mutate(tw = paste(str_to_title(window),
                      str_to_title(stimulus))) %>%
    mutate(tw = factor(tw, levels = c('Pre Cue', 'Post Cue', 'Post Target')),
           task = factor(task, levels = c('Odd/Even', 'Number/Letter')),
           condition = factor(condition, labels = c('Repeat', 'Switch'))) %>%
    select(-stimulus, -window)
  fpath_single_trial_var_meas <- paste(paths$bids,
                                       'derivatives',
                                       'analysis_dataframes',
                                       paste0(meas, '_single_trial.rds'),
                                       sep = '/')
  saveRDS(meas_df,
          file = fpath_single_trial_var_meas)
}
```
